Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-06-29 code/rmarkdown/article.Rmd
+ article fixes

## Introduction
<!--  main text -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritizations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritizations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg. connectivity between populations). For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions (ie. the \textit{BLM} parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490; but see @r489 for alternative methods]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]--especially since most large-scale exercises involve hundreds of species. Thus prioritizations may fail to secure a sufficient proportion of the intra-specific genetic variation of  species, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have been investigating the use of surrogate data to generate prioritizations that adequately preserve intra-specific genetic variation--but without actually needing to utilize genetic data to acheive this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritizations that would preserve more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Surrogates for neutral variation have been based on variables that predict the level of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively). Although these variables may correlate with genetic variation, they may not be effective surrogates if they have poor predictive ability or if these correlations are driven by spatial auto-correlation. Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively). We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations. We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect. If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

We chose this data set for several reasons. First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories. To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species. The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies. Second, we required genetic data for all species to be collected using a standardized sampling scheme, and this data set used a grid to collect samples. Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables. This data set spans across large attitudinal gradients. Although this data set contains AFLPs  which not ideal as genomic representatives (cf. single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists. Thus this data set was well suited for our study.

This data set has previously been used to address a range of different questions. Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486]. However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning. Taberlet _et al._ [-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species. They found that the prioritization based on species distributions was substantially different to the prioritization based on genomic data. Their results suggest that prioritizations based on species distribution data alone will be unable to secure intra-specific genetic variation. However, whether or not environmental and geographic variables can augment prioritizations remains untested.

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilized the sampling grid used by the IntraBioDiv project as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation. We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

@@

The previous set of analyses resulted in an adaptive (if detected) and neutral genetic space for each species. For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space. Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci. Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space. By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].

## Prioritization method
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance. Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

@@

To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations and computed the proportion of intra-specific genetic variation they secured. For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritizations by randomly selecting different combinations of planning units that were occupied by the species. We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritization. We then generated another random set of prioritizations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritization.

We fit two full generalized linear models (GLM) with a logit links. The first model was fit to the proportion of adaptive genetic variation secured in a prioritization, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritization, a categorical variable indicating the species for which the prioritization was generated, and an interaction term between these predictor variables. The second model was fit to the proportion of neutral genetic variation secured in a prioritization, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritization. Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritizations were generated, and an interaction term. We subjected these models to a backwards step-wise term deletion routine to assess term significance. To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively). We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].

To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. However, most large-scale prioritizations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises (eg. by \texttt{Marxan}). We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species. To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations. We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritizations.

We investigated the benefits of using surrogate-based targets in prioritizations using generalized linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469]. We fit a full model to the proportion of genetic variation secured in a given prioritization. This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritizations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritization (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables. This model accommodated prioritizations generated for the same species using a random intercept term. We subject the full model to a backwards step-wise term deletion routine to assess term significance. We conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].

## Results
<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 26 species. The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species. Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (Table S2). On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial auto-correlation--especially the adaptive genetic spaces (eg. _Cerastium uniflorum_ and _Dryas octopetala_). Additionally, several species show east-west patterns in their intra-specific genetic variation (eg. _Peucedanum ostruthium_ and _Phyteuma betonicifolium_).

<!-- correlation analysis -->
The dissimilarity matrices generated using the environmental and geographic variables correlated with those matrices based on the genetic characteristics for most species (Table 1). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics. After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.

>>>>

## Introduction
<!--  main text -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg. connectivity between populations). For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalise fragmented solutions (ie. the \textit{BLM} parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this variation [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490; but see @r489 for alternative methods]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]--especially since most large-scale exercises involve hundreds of species. Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of  species, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have been investigating the use of surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--but without actually needing to utilise genetic data to acheive this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritisations that would preserve more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Surrogates for neutral variation have been based on variables that predict the level of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively). Although these variables may correlate with genetic variation, they may not be effective surrogates if they have poor predictive ability or if these correlations are driven by spatial auto-correlation. Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively). We expect that prioritisations generated using surrogate-based targets will secure more genetic variation than prioritisations generated using only amount-based targets with the same number of planning units. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations. We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect. If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

We chose this data set for several reasons. First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories. To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species. The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies. Second, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples. Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables. This data set spans across large attitudinal gradients. Although this data set contains AFLPs  which not ideal as genomic representatives (cf. single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists. Thus this data set was well suited for our study.

This data set has previously been used to address a range of different questions. Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486]. However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning. Taberlet _et al._ [-@r486] investigated whether prioritisations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species. They found that the prioritisation based on species distributions was substantially different to the prioritisation based on genomic data. Their results suggest that prioritisations based on species distribution data alone will be unable to secure intra-specific genetic variation. However, whether or not environmental and geographic variables can augment prioritisations remains untested.

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation. We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

@@

The previous set of analyses resulted in an adaptive (if detected) and neutral genetic space for each species. For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space. Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci. Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space. By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation [see @r465 for discussion on attribute spaces].

## Prioritisation method
We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance. Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

@@

To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured. For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species. We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation. We then generated another random set of prioritisations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritisation.

We fit two full generalised linear models (GLM) with a logit links. The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritisation, a categorical variable indicating the species for which the prioritisation was generated, and an interaction term between these predictor variables. The second model was fit to the proportion of neutral genetic variation secured in a prioritisation, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritisation. Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritisations were generated, and an interaction term. We subjected these models to a backwards step-wise term deletion routine to assess term significance. To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritisation and the environmental and geographic variation they also secured (respectively). We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].

To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritisations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritisations for each species. However, most large-scale prioritisations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritisations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritisations. Finally, most large-scale prioritisations also explicitly consider opportunity cost to ensure that prioritisations are cost-effective. Our third scenario involved generating multi-species prioritisations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritisations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritisations served as our baseline--representing the prioritisations typically generated in conservation planning exercises (eg. by \texttt{Marxan}). We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species. Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species. To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations. We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.

We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469]. We fit a full model to the proportion of genetic variation secured in a given prioritisation. This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritisations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritisation (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables. This model accommodated prioritisations generated for the same species using a random intercept term. We subject the full model to a backwards step-wise term deletion routine to assess term significance. We conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].

## Results
<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species. The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species. Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(adaptive.k)`$; Table S2). On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg. _Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.

<!-- correlation analysis -->
The dissimilarity matrices based on the environmental and geographic characteristics of the planning units correlated with the dissimilarity matrices based on the genetic characteristics of individuals in the planning units (Table 1). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). These results suggest that prioritisations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics. After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritisation and the proportion of environment variation it also secured was found to depend on the species for which the prioritisation was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). Similarly, the proportion of neutral genetic variation in a prioritisation was affected by the interaction between species and the proportion of neutral variation secured in the prioritisation (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). After establishing a relationship between the proportion of the surrogate variables and proportion of genetic variation secured in a prioritisation, the next step was to determine if prioritisations benefited from the addition of surrogate-based targets and verify if such benefits were evident across a variety of planning scenarios.
